---
title: "Lab1"
author: "Jack Schwarz"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RColorBrewer)
library(ggplot2)
```


# Tasks

## Task 1

Find the working directory

```{r}
library(ggplot2)
```

## Task 2

Read in DDT Data

```{r}
ddt <- read.csv("DDT.csv")
head(ddt)
```

## Task 3

  a.    The Qualitative variables are RIVER and SPECIES
  b.    The Quantitative variables are MILE, LENGTH, WEIGHT, and DDT
  c.    There are 3 different species of fish in the data set.
```{r}
spc <- as.vector(as.matrix(ddt['SPECIES']))
length(unique(spc))
```
  
  d.    Below are the observations with SPECIES = LMBASS and Weight > 800.
```{r}
with(ddt, ddt[WEIGHT>800 & SPECIES=="LMBASS",])
```
  e.    Below are the observations when RIVER = SCM and DDT > 4.0.
```{r}
with(ddt, ddt[DDT > 4.0 & RIVER=="SCM",])
```

## Clicker Questions
  1. The Avg length of fish is: `r mean(ddt[,'LENGTH'])`. (`mean(ddt[,'LENGTH'])`)
  2. The standard deviation of the weight of the fish is: `r sd(ddt[,'WEIGHT'])`. (`sd(ddt[,'WEIGHT'])`)
  3. **No**, that is not the correct Graph.
```{r, fig.align='center'}
ggplot(ddt, aes(x=WEIGHT, y=LENGTH)) + geom_point() + ggtitle("Jack Schwarz")
```
  4.    The last value of v=1:20 is 1.0   
```{r}
v=1:20
v/20
```
  
## Task 4
  a.    Below is the River Table
```{r}
rt=with(ddt,table(RIVER))
rt
```
  b.    Here is a barplot of the above table
```{r, fig.align='center'}
colr = brewer.pal(4, "Set3")
barplot(rt,legend=TRUE, xlim = c(0,5.7),col=colr, xlab="Rivers", main="Number of Rivers")
```
  c.    Below is a cross table of Rivers and Species of Fish
```{r}
dtable = with(ddt,table(SPECIES,RIVER))
dtable
```
  d.    Here is the barplot of the above table
```{r,fig.align='center'}
colr2= brewer.pal(3, "Set3")
barplot(dtable,beside=TRUE,legend=TRUE,col=colr2, xlab="Rivers", ylab="Number of Fish",main="Species of Fish by River",xlim = c(0,23))
```

## Task 5

  a. Below is the Species pie chart.
```{r,fig.align='center'}
sp=with(ddt,table(SPECIES))
pie(sp,col=3:7)
```

  b.  Here is the River Pie Chart
```{r}
pie(rt, col=2:6)
```

## Task 6
Below are a, b, c. These are the boxplots of DDT, Weight, and Length.
```{r, align='center'}
layout(matrix(c(1,2,3),nr=1,nc=3))
with(ddt,boxplot(DDT,ylab="DDT",col="Blue", main="DDT",notch=TRUE))
with(ddt,boxplot(WEIGHT,ylab="WEIGHT",col="Green", main="Weight",notch=TRUE))
with(ddt,boxplot(LENGTH,ylab="LENGTH",col="Red", main="Length",notch=TRUE))
```

## Task 7

  a.  Here is the coplot of Length Vs Weight given River
```{r, align='center'}
rivcol=with(ddt, ifelse(RIVER=="FCM","Red",
                        ifelse(RIVER=="LCM","Blue",
                               ifelse(RIVER=="SCM","Green","Black"))))

coplot(LENGTH~WEIGHT|RIVER,data=ddt,col=rivcol)
```
  
  b.    Here is the coplot of DDT Vs Weight given Species
```{r, align='center'}
fishcol=with(ddt,ifelse(SPECIES=="CCATFISH","Red",
                        ifelse(SPECIES=="SMBUFFALO","Blue","Green")))

coplot(DDT~WEIGHT|SPECIES,data=ddt,col=fishcol)
```
  

## Task 8
  a. Here is the ggplot Boxplot
```{r, align='center'}
ggplot(ddt, aes(x=SPECIES, y=WEIGHT, fill=RIVER)) +
  geom_boxplot() + ggtitle("Jack Schwarz")
```

  b.    Here is the ggplot Violin Plot
```{r, align='center'}
ggplot(ddt, aes(x=RIVER, y=LENGTH, fill=SPECIES)) +
  geom_violin() + ggtitle("Jack Schwarz")
```

  c.    Here is the ggplot Scatter Plot
```{r, align='center'}
ggplot(ddt, aes(x=WEIGHT, y=LENGTH, fill=SPECIES, color=SPECIES)) +
  geom_point() + ggtitle("Jack Schwarz")
```
